The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies

### 2016-08-16 article/unrave.tex
I should really spent *all* of that time writing..

% Look for \stub{}s. If you can expand it, do so!

\documentclass[preprint2,trackchanges]{aastex}

@@

@@

\begin{document}

\title{The \project{unRAVE} catalog}

@@

\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W. Hogg, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any other \project{RAVE} people who have contributed to the data analysis}
%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}

@@

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for approximately 290,000 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
that employs prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
\end{abstract}

@@

@@

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> Asteroseismic nu_max and delta_nu available anywhere?
% --> Literature studies presented in Kordopatis+ 2014

@@

\acknowledgements
We thank Sven Buder (MPIA), others??.
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

@@

@@

@@

@@

@@

@@

>>>>

%% TODO:
%% Co-authors: Look for \stub{}s. If you can expand it, do so!

\documentclass[preprint,trackchanges]{aastex}

@@

% This section can be removed at submission.
% ------------------------------------------
\usepackage{color}
\usepackage{datenumber}

\newcounter{dateone}
\newcounter{datetwo}

\newcommand{\difftoday}[3]{%
  \setmydatenumber{dateone}{\the\year}{\the\month}{\the\day}%
  \setmydatenumber{datetwo}{#1}{#2}{#3}%
  \addtocounter{datetwo}{-\thedateone}%
  \the\numexpr(\thedatetwo)\relax\space days
}
% ------------------------------------------

@@

\begin{document}
% Remove at submission:
\slugcomment{{\color{red} \textbf{To appear on arXiv on 9 September 2016 (\difftoday{2016}{9}{7}away)}}}

\title{The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies}

@@

\author{Some combination of: Harry Enke, Gerry Gilmore, Keith Hawkins, David W. Hogg, Georges Kordopatis, Gal Matijevic, Melissa Ness, Jason Sanders, Matthias Steinmetz, Hans Walter-Rix, and any \project{RAVE} members who have contributed to the data analysis in this study}
%\altaffiltext{3}{\texttt{arc@ast.cam.ac.uk}}

@@

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for approximately 290,000 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
that employs prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
% For stars not in TGAS, we calculate distances
\end{abstract}

@@

The Milky Way is considered to be the best laboratory for understanding galaxy formation and evolution.  This premise depends on the ability to precisely measure the position, motion, and chemistry of (many) individual stars, and to infer the structure, kinematics, and chemical enrichment of the Galaxy from those data \citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,Jofre_2015,Madler_2016}, proper motions are frequently plagued by unquantified systematics from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and stellar spectroscopists frequently report significantly different chemical abundance patterns \citep{Smiljanic_2014}.  If these measurements were known with high fidelity then \stub{xyz}.

% These scientific questions were only some of them that prompted Gaia.
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

% Gaia will produce a very clear snapshot, producing astrometry space motions

% but if we want to infer the past then we need additional information about how the galaxy has been enriched. so we need chemical abundances.
% the number of stars that overlap with Gaia DR1 between stellar surveys is very different, due to targeting, etc
% and those stars are not on the same metallicity scale
% therefore a large-scale catalog of homogenous abundances that overlap with Gaia DR1 is currently lacking

@@

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> APOGEE / EPIC (whichever was not used)
% --> Literature studies presented in Kordopatis+ 2014

@@

\acknowledgements
We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), others??.
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

@@

\bibitem[Casey et al.(2012)]{Casey_2012} Casey, A.~R., Keller, S.~C., \& Da Costa, G.\ 2012, \aj, 143, 88

\bibitem[Casey et al.(2013)]{Casey_2013} Casey, A.~R., Da Costa, G., Keller, S.~C., \& Maunder, E.\ 2013, \apj, 764, 39

\bibitem[Casey et al.(2014a)]{Casey_2014a} Casey, A.~R., Keller, S.~C., Da Costa, G., Frebel, A., \& Maunder, E.\ 2014, \apj, 784, 19

\bibitem[Casey et al.(2014b)]{Casey_2014b} Casey, A.~R., Keller, S.~C., Alves-Brito, A., et al.\ 2014, \mnras, 443, 828

\bibitem[Casey \& Schlaufman(2015)]{Casey_Schlaufman_2015} Casey, A.~R., \& Schlaufman, K.~C.\ 2015, \apj, 809, 110

\bibitem[Casey et al.(2016)]{Casey_2016} Casey, A.~R., Hogg, D.~W., Ness, M., et al.\ 2016, arXiv:1603.03040

@@

\bibitem[Deason et al.(2011)]{Deason_2011} Deason, A.~J., Belokurov, V., \& Evans, N.~W.\ 2011, \mnras, 416, 2903

@@

\bibitem[Jofr{\'e} et al.(2015)]{Jofre_2015} Jofr{\'e}, P., M{\"a}dler, T., Gilmore, G., et al.\ 2015, \mnras, 453, 1428

@@

\bibitem[M{\"a}dler et al.(2016)]{Madler_2016} M{\"a}dler, T., Jofr{\'e}, P., Gilmore, G., et al.\ 2016, arXiv:1606.03015

\bibitem[Ness et al.(2012)]{Ness_2012} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2012, \apj, 756, 22

\bibitem[Ness et al.(2013a)]{Ness_2013a} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 430, 836

\bibitem[Ness et al.(2013b)]{Ness_2013b} Ness, M., Freeman, K., Athanassoula, E., et al.\ 2013, \mnras, 432, 2092

@@

\bibitem[Smiljanic et al.(2014)]{Smiljanic_2014} Smiljanic, R., Korn, A.~J., Bergemann, M., et al.\ 2014, \aap, 570, A122

@@

\bibitem[van Leeuwen(2007)]{van_Leeuwen_2007} van Leeuwen, F.\ 2007, \aap, 474, 653
